Altering Backward Pass Gradients improves Convergence

### 2022-09-16 paper.tex
Modifications 5.

This section empirically examines some of the issues that occur when networks are
trained without normalization layers. ResNet-18 \cite{he2016deep} is used as the
foundation model for all empirical trials owing to its popularity among deep learning
practitioners and the relative simplicity with which it can be trained on a big dataset
such as ImageNet \cite{russakovsky2015imagenet}. ResNet-18 is composed of $20$
convolutional layers and one fully connected (FC) layer. At the input, it has a single
convolutional layer with filters of size $7\times 7$. Three convolutional layers are
placed in the downsampling skip connections. The remaining convolutional layers have
several $3\times 3$ filters that are grouped together to create a block. Each basic
block is connected by a skip connection. At the end of all convolutional layers, a
global average pooling layer downsamples the features to a fixed-size feature vector and
transfers it to the FC layer, which fits the features using regression and outputs
logits. The logits $z_i$ produced by the FC layer are fed into a softmax layer, which
transforms them to predicted probabilities $p_i$ using the equation $p_i=\frac
{e^{z_i}}{\sum _i e^{z_i}}$.

We concentrate on the following metric: the per-filter L2-norm of each layer. Throughout
the training process, we monitor variations in the per-filter weight norm. Because a
full plot of each layer's data would use an inordinate amount of space, we focus on a
few critical layers' plots. The supplementary section contains plots for each layer.
ResNet-18 without any normalization layers is used for all experiments in this section.
The norm of the weights of the $11\textit{th}$ convolutional layer is shown in Fig.
\ref{fig:norm_plots}a. Layer $11$ includes $256$ filters. Each colour on the figure
represents the evolution of a particular filter throughout the course of training
iterations. The critical point to note is that some filters achieve a norm of zero
during training. This event is referred to as `Zeroing Out', and occurs when a channel
of a weight tensor gets fully filled with zeros. If all the coefficients of the filter
is 0 then only L2 norm of a filter can become zero. If we observe zero filter norm for a
filter, then that filter does not contribute at all to determine the input-output
relationship of a dataset. According to our findings, when a filter tensor becomes
zeroed out, it does not recover with further training.

@@

In a similar fashion to the study in section \ref{sec:Empi}, we examine the norm
profiles of features and weights in the PGT and AGC. The supplementary section includes
all the detailed graphs of each layer. In Fig. \ref{fig:norm_plots}d, we show the
per-filter norm plot of the $11\textit{th}$ convolutional layer as it progresses through
a PGT enabled training session. We observe that the number of zeroed out filters has
considerably reduced. We observe similar benefits in plot of Fig. \ref{fig:norm_plots}e,
which is the plot of the per-filter norm of the final convolutional layer. Coming to
benefits in feature norms, we observe in plot of Fig. \ref{fig:norm_plots}f, the final
feature tensor obtained with PGT enabled training do not contain any zeroed out regions,
leading us to assert that information loss is heavily mitigated as the features pass on
from the feature extracting convolutional layers to the fully connected stage for
regression.

AGC is excellent at eliminating the issue of filter and feature zeroing out. However,
AGC recovers a lower performance benefit than PGT. In the instance of PGT, we see that
it also mitigates the zeroing out phenomenon in most layers. In Fig.\ref{fig:agc_pgt}a,
we see a plot of the $15\textit{th}$ layer of the non-BN ResNet-18 trained without PGT.
As we can see that there has been quite a few filter dropouts in the baseline. PGT
recovers most of the filters [Fig. \ref{fig:agc_pgt}b]. In some layers however, the
performance of PGT worsens. The $2\textit{nd}$ convolutional layer has more zeroed-out
filters in PGT [Fig. \ref{fig:agc_pgt}e] than in the baseline [Fig. \ref{fig:agc_pgt}d].
Despite that, PGT achieves a greater overall accuracy. This necessitates the
simultaneous use of AGC and PGT.

>>>>

This section empirically examines some of the issues that occur when networks are
trained without normalization layers. ResNet-18 \cite{he2016deep} is used as the
foundation model for all empirical trials owing to its popularity among deep learning
practitioners and the relative simplicity with which it can be trained on a big dataset
such as ImageNet \cite{russakovsky2015imagenet}. ResNet-18 is composed of $20$
convolutional layers and one fully connected (FC) layer. At the input, it has a single
convolutional layer with filters of size $7\times 7$. Three convolutional layers are
placed in the downsampling skip connections. The remaining convolutional layers have
several $3\times 3$ filters that are grouped together to create a block. Each basic
block is connected by a skip connection. At the end of all convolutional layers, a
global average pooling layer downsamples the features to a fixed-size feature vector and
transfers it to the FC layer, which fits the features using regression and outputs
logits. The logits $z_i$ produced by the FC layer are fed into a softmax layer, which
transforms them to predicted probabilities $p_i$ using the equation $p_i=\frac
{e^{z_i}}{\sum _j e^{z_j}}$.

We concentrate on the following metric: the per-filter L2-norm of each layer. Throughout
the training process, we monitor variations in the per-filter weight norm. Because a
full plot of each layer's data would use an inordinate amount of space, we focus on a
few critical layers' plots. The supplementary section contains plots for each layer.
ResNet-18 without any normalization layers is used for all experiments in this section.
The norm of the weights of the $11^\textit{th}$ convolutional layer is shown in Fig.
\ref{fig:norm_plots}a. Layer $11$ includes $256$ filters. Each colour on the figure
represents the evolution of a particular filter throughout the course of training
iterations. The critical point to note is that some filters achieve a norm of zero
during training. This event is referred to as `Zeroing Out', and occurs when a channel
of a weight tensor gets fully filled with zeros. If all the coefficients of the filter
is 0 then only L2 norm of a filter can become zero. If we observe zero filter norm for a
filter, then that filter does not contribute at all to determine the input-output
relationship of a dataset. According to our findings, when a filter tensor becomes
zeroed out, it does not recover with further training.

@@

In a similar fashion to the study in section \ref{sec:Empi}, we examine the norm
profiles of features and weights in the PGT and AGC. The supplementary section includes
all the detailed graphs of each layer. In Fig. \ref{fig:norm_plots}d, we show the
per-filter norm plot of the $11^\textit{th}$ convolutional layer as it progresses through
a PGT enabled training session. We observe that the number of zeroed out filters has
considerably reduced. We observe similar benefits in plot of Fig. \ref{fig:norm_plots}e,
which is the plot of the per-filter norm of the final convolutional layer. Coming to
benefits in feature norms, we observe in plot of Fig. \ref{fig:norm_plots}f, the final
feature tensor obtained with PGT enabled training do not contain any zeroed out regions,
leading us to assert that information loss is heavily mitigated as the features pass on
from the feature extracting convolutional layers to the fully connected stage for
regression.

AGC is excellent at eliminating the issue of filter and feature zeroing out. However,
AGC recovers a lower performance benefit than PGT. In the instance of PGT, we see that
it also mitigates the zeroing out phenomenon in most layers. In Fig.\ref{fig:agc_pgt}a,
we see a plot of the $15^\textit{th}$ layer of the non-BN ResNet-18 trained without PGT.
As we can see that there has been quite a few filter dropouts in the baseline. PGT
recovers most of the filters [Fig. \ref{fig:agc_pgt}b]. In some layers however, the
performance of PGT worsens. The $2\textit{nd}$ convolutional layer has more zeroed-out
filters in PGT [Fig. \ref{fig:agc_pgt}e] than in the baseline [Fig. \ref{fig:agc_pgt}d].
Despite that, PGT achieves a greater overall accuracy. This necessitates the
simultaneous use of AGC and PGT.
